Living on the edge: crop boundary effects on precision canola and wheat yields in Alberta

### 2021-09-29 Manuscript/manuscript.Rmd
Updated figures

Boundary effects are difficult to observe without large amounts of crop yield data, which is typically expensive and time-consuming to collect.    
Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
Meta-analyses are a common solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs. 
However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems [@redhead2020]; typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.

@@

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013].
We filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time.

@@

Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the data.
Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.      <!-- - such as strip farming -->
Ground speed is extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields.
Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
To model this, we fit the following model to each field-year of data:

  \begin{equation}
  \begin{split}
  \sqrt{yield} \sim & Normal (\mu, \sigma)\\
  \mu = & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
  log(\sigma) =  & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
  \end{split}
  \end{equation}

@@

Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation.
_Not really sure how we can get around this. In practice, this means that SEs for all coefficients are too small (i.e. we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points)._

<!-- - This is extremely common in precision yield data, but  -->

@@

SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (e.g. Lan's paper) which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021], but this is not broadly studied.
Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat [but see @fahrig2019].

>>>>

Boundary effects are difficult to observe without large amounts of crop yield data, which is typically expensive and time-consuming to collect.    
Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
Meta-analyses are a common solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs. 
However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems [@redhead2020; @hunicken2021]; typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.

@@

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m$^2$), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013], or when the combine is accelerating or decelerating [@arslan2002].
Therefore, we used the following filtering procedure to clean the data prior to model fitting:

1. Filtered spatial outliers and "inliers" using the procedure from [@vega2019] 
2. Filtered unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), or values less than or equal to zero
3. Filtered extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
4. Filtered points with large changes (>20%) in ground speed
5. Filtered extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between points

Due to the large number of data at each field, we sub-sampled filtered data to 50,000 data points per field to reduce computation time.

@@

Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.      <!-- - such as strip farming -->
Ground speed is extremely important to yield monitor accuracy, with low ground speed registering higher yields.
Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
To model this, we fit the following model to each field-year of data:

  \begin{equation}
  \begin{split}
  \sqrt{yield} \sim & Normal (\mu, \sigma)\\
  \mu = & Intercept + f(\text{Edge Distance}, b=5)_i + \\
   & f(\text{Easting}, \text{Northing}, b=60)\\
  log(\sigma) =  & Intercept + f(\text{Edge Distance}, b=5)_i + \\
   & f(\text{Easting}, \text{Northing}, b=60)\\
  \end{split}
  \end{equation}

@@

The residuals in most of the models displayed both spatial and temporal autocorrelation in the residuals, representing "micro-scale" variation [@cr] that was not captured by the spatial smoothers.

This is extremely common in precision yield data.

We did not include temporal smoothers because spatial and temporal processes are confounded in our model; that is, combining of fields occurs in patterns (e.g. east-to-west) so it is impossible to tell whether lower or higher yields in certain areas of the field were caused by deviations in the yield monitor sensors over time.
Therefore, we assigned all random variation as spatial in nature; customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.

<!-- _Not really sure how we can get around this. In practice, this means that SEs for all coefficients are too small (i.e. we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points)._ -->

@@

SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (e.g. Lan's paper) which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021; @haslem2021], but this is not broadly studied.
Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat [but see @fahrig2019].
